\begin{abstract}
  Experimental evolution (EE) studies are powerful tools for observing
  molecular evolution ``in-action'' from populations sampled in
  controlled and natural environments. The advent of next generation
  sequencing technologies has made whole-genome and whole-population
  sampling possible, even for eukaryotic organisms with large genomes,
  and allowed us to locate the genes and variants responsible for
  genetic adaptation. While many computational tests have been
  developed for detecting regions under selection, they are mainly
  designed for static (single time) data, and work best when the
  favored allele is close to fixation.
	
  EE studies provide samples over multiple time points, underscoring
  the need for tools that can exploit the data. At the same time, EE
  studies are constrained by the limited time span since onset of
  selection, depending upon the generation time for the organism. This
  constraint impedes adaptation and optimization studies, as the
  population can only be propagated for a small number of generations 
  (depends on organism generation time),
  relative to the fixation time (mainly depends on the strength of selection) of 
  the favored allele. Moreover, coverage of a 
  variant in pool-sequenced (pool-seq)  experiments vary across replicates and
  time points, leading to heterogeneous ascertainment bias in measuring
  population allele frequency across different measurements.
	
  In this article, we directly address these issues while developing
  tools for identifying selective sweep in pool-seq EE of sexual organisms and
  propose Composition of Likelihoods for Evolve-And-Resequence 
  experiments (\comale). Extensive simulations show that
  \comale\ achieves higher power in detecting and localizing selection over a 
  wide range of parameters. 
  In addition, we show that \comale\ consistently estimates model parameters, 
  selection strength and overdominance, as byproduct of the statistical test, 
  while being orders of magnitude faster than existing methods.
  We also demonstrate that while the performance of existing methods are 
  sensitive the sequencing coverage, \comale\ statistics is robust to the 
  variation of coverage. 
  Finally, we apply the \comale\ statistic to the
  study of adaptation of \dmel to the novel laboratory environments 
  (alternating hot and cold temperatures) to identify adaptive genes/alleles. 
  
  \AI{one sentence of real data results}
\end{abstract}