\section{Results}
\paragraph{Modeling Allele Frequency Trajectories in Small Populations.} 
We first tested the goodness of fit of the discrete versus continuous models in 
modeling allele frequency trajectories, under general E\&R parameters.  For 
this 
purpose, we conducted $100$K
simulations with two time samples $\Tc=\{0,\tau\}$ where $\tau\in
\{1,10,100\}$ is the parameter controlling the density of sampling
in time.  In addition, we repeated simulations for different values of
starting frequency $\nu_0\in\{0.005,0.1\}$ (i.e., hard and soft sweep)
and selection strength $s\in\{0,0.1\}$ (i.e., neutral and
selection). Then, given initial frequency $\nu_0$, we computed
the expected distribution of the frequency of the next sample $\nu_\tau$
under two models to make a comparison.  \ref{fig:markov}A-F shows that
Brownian motion (continuous model) is inadequate when $\nu_0$ is
far from $0.5$, or when sampling times are sparse ($\tau>1$). If the
favored allele arises from standing variation in a neutral population,
it is unlikely to have frequency close to $0.5$, and the starting
frequencies are usually much smaller (see
\ref{fig:sfs}). Moreover, in typical \dmel experiments for
example, sampling is sparse. Often, the experiment is designed so that
$10\le\tau\le100$~\cite{kofler2013guide, orozco2012adaptation,
  zhou2011experimental,franssen2015patterns}.

% XXX Not sufficient to say 'in contrast to brownian motion, markov...
% Please check the rewrite of this part
In contrast to the Brownian motion approximation, discrete Markov
chain predictions (Eq.~\ref{eq:mkvs}) are highly consistent with
empirical data for a wide range of simulation parameters
(\ref{fig:markov}A-M). Moreover, the discrete markov chain can be
modified to model the case when the the allele is under
selection. 

\paragraph{Detection Power.} 
We compared the performance of \comale\ against other methods for
detecting selection. For each method we calculated detection power as the 
percentage of true-positives identified with false-positive rate $\le 0.05$. For 
each
configuration (specified with values for selection coefficient $s$,
starting allele frequency $\nu_0$ and coverage $\lambda$), power of each method 
is evaluated over $2000$ distinct simulations, half of which modeled neutral 
evolution
and the rest modeled positive selection.



We compared the power of \comale\ with Gaussian process
(GP)~\cite{Terhorst2015Multi}, FIT~\cite{feder2014Identifying}, and
CMH~\cite{agresti2011categorical} statistics.  FIT and GP convert read counts
to allele frequencies prior to computing the test statistic.  \comale\
shows the highest power in all cases and the power stays relatively
high even for low coverage (\ref{fig:power} and
\ref{tab:power}). In particular, the difference in performance
of \comale\ with other methods is pronounced when starting frequency
is low. 
The
advantage of \comale\ stems from the fact that favored allele with low
starting frequency might be missed by low coverage sequencing. In this
case, incorporating the signal from linked sites becomes increasingly
important. We note that methods using only two time points, such as
CMH, do relatively well for high selection values and high
coverage. However, the use of time-series data can increase detection
power in low coverage experiments or when starting frequency is
low. Moreover, time-series data provide means for estimating selection
parameters $s,h$ (see below). Finally, as \comale\ is robust to change
of coverage, our results (\ref{fig:power}B,C) suggest that taking
many samples with lower coverage is preferable to sparse sampling with
higher coverage.


\ignore{ Moreover, the observation of
  \ref{fig:markov} is revisited here, approximate continues
  models (GP and FIT) perform poorly when starting frequency is low
  (\ref{fig:power}A-C).  

\paragraph{SFS for Detection in Natural Samples.} We did not show the
SFS based statistics in \ref{fig:power} as they did not perform
better than random. In majority of controlled experimental evolution studies, 
the population is restricted set of $F$ founder lines, where $F<<N_e$
(\ref{fig:ee}B) and inbred during the experiment. This creates a severe 
bottleneck, confounding
SFS. ~\ref{fig:bottleneck} demonstrates the effect of
experimental evolution on different SFS statistics under neutral
evolution for 1000 simulations. A second problem with using SFS for
experimental evolution is that the sampling starts right after the
onset of experimentally induced selection, and the favored allele may
not reach high enough frequency to modify the site frequency spectrum
(\ref{fig:sweep}).

However, in experiments involving naturally evolving populations,
even if the span of the time-series is small, the onset of selection
might occur many generations prior to sampling. To test performance of
SFS-based statistics in natural evolution, using \texttt{msms}, we conducted 
200 (100
neutral and 100 sweep) forward simulations for different values of
$s$, $N_e=10K$ and $N=200$. The start
of sampling was chosen randomly after onset of selection in two distinct 
scenarios. Let $t_{\nu=x}(s,N_e)$ denote
the expected time (in generations) required to reach carrier frequency
$x$ in a hard sweep and $U[a,b]$ denote discrete uniform distribution
in the interval $[a,b]$. First we considered the case when start of
sampling is chosen throughout the whole sweep. i.e., $\tau_0 \sim
U\left[1,t_{\nu=1}(s,N_e)\right]$ (\ref{fig:powerSFS}A). Next, we
considered sampling start time chosen nearer to fixation of the
favored allele, i.e., $\tau_0 \sim
U\left[t_{\nu=0.9}(s,N_e),t_{\nu=1}(s,N_e)\right]$
(\ref{fig:powerSFS}B). In both scenarios, sampling was done over
$6$ time points within $50$ generations of $\tau_0$ (~\ref{fig:ee}A). We 
compared
\comale, GP, FIT with both static and dynamic SFS based statistics of
SFSelect and Tajima's D. ~\ref{fig:powerSFS}A shows that SFS based
statistics are outperformed by other methods. However,
when sampling is performed close to fixation, i.e., when the favored
allele has frequency of 0.9 or higher, SFS based statistics perform
considerably better than GP, FIT and \comale\
(\ref{fig:powerSFS}B). Moreover, dynamic SFS statistics
provide higher power than static SFS statistics, demonstrating that in the use 
of time-series SFS based statistics is advantageous.}


\paragraph{Site-identification.}
In general, localizing the favored variant, using pool-seq data is a
nontrivial task due to extensive linkage
disequilibrium~\cite{tobler2014massive}.  To measure performance, we
sorted variants by their $H$ scores and computed rank of the favored
allele for each method. For each setting of $\nu_0$ and $s$, we
conducted $1000$ simulations and computed the rank of the favored
mutation in each simulation. The cumulative distribution of the rank
of the favored allele in 1000 simulation for each setting
(\ref{fig:rank}) shows that \comale\ outperforms other statistics.

An interesting observation is revisiting the contrast between
site-identification and
detection~\cite{long2013massive,tobler2014massive}.  When selection
strength is high, detection is easier (\ref{fig:power}A-F), but
site-identification is harder, due to the high LD between flanking
variants and the favored allele (\ref{fig:rank}A-F).  Moreover,
site-identification becomes more difficult whenever the initial
frequency of the favored allele is low, i.e., at the onset of
selection, LD between favored allele and its nearby variants is
high. For example, when coverage $\lambda=100$ and selection
coefficient $s=0.1$, the detection power is 75\% for hard sweep, but
100\% for soft sweep (\ref{fig:power}B-E). In contrast, the favored
site was ranked as the top in 14\% of hard sweep cases, compared to
and 95\% of soft sweep simulations.  \ignore{
  \newcommand*\rot{\rotatebox{-90} }
\begin{table}[H]
	\centering
		\caption{\bf Percentage of simulations which favored allele appears in 
		top of the ranking.}
	\begin{tabular}{c||c}
		 Hard Sweep &Soft Sweep\\
		\hline
		(A) & (B)  \\
		{\input{../tables/top1.100.0.hard.tex}}
		 &{\input{../tables/top1.100.0.soft.tex} }\\
		\hline
		(C)  & (D)  \\
		{\input{../tables/top10.100.0.hard.tex}}
		& {\input{../tables/top10.100.0.soft.tex} }\\
		\hline
		(E) & (F) \\
		{\input{../tables/top50.100.0.hard.tex}}
		&{\input{../tables/top50.100.0.soft.tex} }\\
		\hline
	\end{tabular}\\
	\caption*{		Percentage of simulations in which the favored allele is 
	ranked 
	first 
		(A-B); 
		appears in top 10 (C-D); or,  appears in top 50 (E-F). In soft sweep 
		simulations (B,D,F), the ranks are consistently better than 
		hard sweep simulations (A,C,E). This can be attributed to lower LD 
		between the 
		hitchhikers (false positives) and favored allele in soft sweep 
		scenarios.}\label{tab:rank}
\end{table}
} 
\paragraph{Estimating Parameters.}
\comale\ estimates effective population size $\hN$  and selection parameters, 
$\hat{s}$ and $\hat{h}$, as
a byproduct of the hypothesis testing. We computed bias of selection
fitness ($s-\hat{s}$) and dominance ($h-\hat{h}$) for of \comale\
and GP for 1000 simulations in each setting. The distribution of the error 
(bias) for
100$\times$ coverage is presented in \ref{fig:bias100} for
different configurations.
\ref{fig:bias30} and \ref{fig:biasinf} provide the
distribution of estimation errors for 30$\times$, and 300$\times$
coverage, respectively.  For hard sweep, \comale\ provides estimates
of $s$ with lower variance of bias (\ref{fig:bias100}A). In soft
sweep, GP and \comale\ both provide unbiased estimates of $s$ with low
variance (\ref{fig:bias100}B). \ref{fig:bias100}~C-D shows
that \comale\ provides unbiased estimates of $h$ as well when 
$h\in\{0,0.5,1,2\}$ and $s=0.1$.
We also tested if \comale\ provide unbiased estimates of $N$, by estimating 
population size on 1000 simulations when $N\in \{200,600,1000\}$. As 
shown in~\ref{fig:estimateN}A-C, maximum likelihood is attained at true value 
of the parameter.

\paragraph{Running Time.}
As \comale\ does not compute exact likelihood of a region (i.e., does
not explicitly model linkage between sites), the complexity of
scanning a genome is linear in number of polymorphisms.  Calculating
score of each variant requires and $\Oc(TRN^3)$ computation
for $\Hc$. However, most of the operations
are can be vectorized for all replicates to make the effective running
time for each variant.  We
conducted $1000$ simulations and measured running times for computing site 
statistics $H$, FIT, CMH and GP with different number of linked-loci.  Our
analysis reveals (\ref{fig:runTime}) that \comale\ is orders of
magnitude faster than GP, and comparable to FIT. While slower than CMH
on the time per variant, the actual running times are comparable after
vectorization and broadcasting over variants (see below).

These times can have a practical consequence. For instance, to run GP
in the single locus mode on the entire pool-seq data of the \dmel genome from a
small sample ($\approx$1.6M variant sites), it would take 1444 CPU-hours
($\approx$ 1 CPU-month). In contrast, after vectorizing and
broadcasting operations for all variants operations using
\texttt{numba} package, \comale\ took 75 minutes to perform an
scan, including precomputation, while the fastest method, CMH, took 17 minutes.

\subsection{Analysis of a \dmel Adaptation to Alternating 
Temperatures}\label{sec:dmel}
We applied \comale\ to the 
\datadm~\cite{orozco2012adaptation,franssen2015patterns}, where
3 replicate samples were chosen from a population of \dmel for 59
generations under alternating 12-hour cycles of  hot stressful (28$^{\circ}$C)
and non-stressful (18$^{\circ}$C) temperatures and sequenced.  In this dataset,
sequencing coverage is different across replicates and generations
(see S2 Fig of~\cite{Terhorst2015Multi}) which makes variant depths
highly heterogeneous (\ref{fig:depthHetero}). 

We first filtered out heterochromatic, centromeric and telomeric
regions~\cite{fiston2010drosophila}, and those variants that have
collective coverage of more that 1500 in all 13 populations: three
replicates at the base population, two replicates at generation 15,
one replicate at generation 23, one replicate at generation 27, three
replicates at generation 37 and three replicates at generation
59. After filtering, we ended up with 1,605,714 variants.

Next, we estimated genome-wide population size $\hat{N}=250$
(\ref{fig:estimateN}-E) which is consistent with previous
studies~\cite{orozco2012adaptation,jonas2016estimating}. The
likelihood curves of \comale\ are sharper around the optimum compared
to that of Bollback et. al~\cite{bollback2008estimation}'s method (see
Supplementary Fig. 1 in~\cite{orozco2012adaptation}).  Also,
chromosomes 3L and 3R appear to have smaller population
size~\ref{fig:estimateN}-D, $\hN=200,150$, respectively. Others have
made similar observations on this data. In particular,
J\'{o}n\'{a}s~\emph{et al.} shown that the chromosome-wise population
size varies even more when it is computed for each replicate
separately (see Table 1 in~\cite{jonas2016estimating}). For instance,
$\hN$ is 131 for chr3R replicate 1, while it is 328 for chrX replicate
2.  While it would be ideal to compute \comale\ statistic for each
replicate and chromosome separately, computing empirical $p$-values
and significant regains become computationally intensive, and
empirical null distribution of each replicate and each chromosome
needs to be computed.  Hence, we use genome-wide estimate $\hN=250$ to
compute $\Hc$ and then compute normalized statistics $\Hc^*$ for each
chromosome, separately.  Specifically, given $\hat{N}=250$, we
computed estimates of $s$, and computed the normalized test statistic
$\Hc^*$ on sliding windows of size of 25Kbp and step size of 5Kbp
variants over the genome.  Empirical null distribution of $\Hc^*$ is
estimated by creating 100 simulations according to
Section~\ref{sec:sims} with $L=20$Mbp, $\hN=250$ and experimental data
coverage and sampling times.  Then, $p$-value of each region in the
experimental data is calculated as the fraction of the 400K null
statistics that are greater than or equal of test
statistic(see~\ref{fig:null-alt}).  After correcting for multiple
testing, we identify $5$ contiguous
intervals~\ref{fig:man-dmel-region} satisfying FDR$\le0.05$, and
covering $2,829$ polymorphic sites. We further performed single-locus
hypothesis testing on the $2,829$ to identify selected variants.
$p$-values of $H$ statistics calculated on a set of 90M single locus
Wright-Fisher simulations for $\widehat{N}=$250, initial frequencies
and variant depths of real data (see \ref{proc:arya}).  $174$ variants
with FDR $\le0.01$\% within selected regions using single locus
hypothesis testing.  The selected 174 variants fall within 32
genes(\ref{tab:genes}) including many Serine inhibitory proteases
(serpins), and other genes involved in endocytosis. Recycling of
synaptic vesicles is seen to be blocked at high temperature in
temperature sensitive Drosophila
mutants~\cite{kosaka1983reversible}. This is also supported by GO
enrichment analysis, where a single GO term `inhibition of
proteolysis' is found to enriched (corrected $p$-value:0.0041).  To
test for dominant selection, we computed $D$ statistic on simulated
neutral and experimental data, and computed $p$-values accordingly.
After correcting for multiple testing, 96 variants were discovered
with FDR$\le 0.01$~(\ref{fig:man-dmel-snp}).

\ignore{
Here is 
\href{https://genome.ucsc.edu/cgi-bin/hgTracks?hgS_doOtherUser=submit&hgS_otherUserName=airanmehr&hgS_otherUserSessionName=Dmel\%20EE\%20HotCold}{UCSC
 track} for this data.
}

\subsection{Analysis of Outcrossing Yeast Populations}
We also applied \comale\ to outcrossing $12$ replicate samples of
Yeast populations~\cite{burke2014standing}, where samples are taken at
generations $\Tc=\{0,180,360,540\}$. We observed a significant
variation in the genome-wide site frequency spectrum of certain
populations over different time points for some
replicates~(\ref{fig:yeast-sfs}). The variation does not have an
easily identifiable cause. Therefore, we focused analysis on seven
replicates $r\in\{3,7,8,9,10,11,12\}$ with genome-wide site-frequency
spectrum over the time range~(\ref{fig:yeast-pca}).

We estimated population size to be $\hN=2000$ haplotypes, and computed
$\hs$, $\hh$ and $H$ statistic accordingly. To compute $p$-values, we
created 1M single-locus neutral simulations according to experimental
data's initial frequency and coverage. By setting FDR cutoff to
$0.05$, only 18 and 16 variants show significant signal for
directional and dominant selection,
respectively~(\ref{fig:man-dmel-snp}).  Selected variants for
directional selection are clustered in two regions, which match $2$ of
the $5$ regions (regions C and E in Fig. 2-a
in~\cite{burke2014standing}) identified by Burke~\emph{et al.} in
their preliminary analysis.  \ignore{ Here is
  \href{https://genome.ucsc.edu/cgi-bin/hgTracks?hgS_doOtherUser=submit\&hgS_otherUserName=airanmehr\&hgS_otherUserSessionName=Yeast\%20EE}{UCSC
    track} for this data.  }
