\documentclass[11pt]{article}
\title{\comale: Composition of Likelihoods for Evolve And Resequence 
Experiments}
\input{../manuscript/macros_arya.tex}
\date{\today}
\usepackage[margin=1in]{geometry} %not compatible with PLOS style
\renewcommand{\figurename}{Fig.}
\def\genetics{\text{\sc{Genetics}}}
\begin{document}
\noindent Dear Editor\\
\noindent\genetics\\

\noindent We would like to submit our manuscript  

``\comale: Composition of Likelihoods for Evolve And Resequence Experiments''

\noindent for possible publication in \genetics.
Our manuscript develops a new method identifying selection in experimental 
evolution studies, which provide an important platform for understanding 
evolution in real-time. 
Specifically, by evolving-and-resequencing a population of an organism, such 
studies 
can help explain how organisms genetically adapt to 
selection pressures such as pesticide, drug, antibiotic resistance. Over past 
decade, there has been a substantial effort has focused on analyzing 
longitudinal studies of sexually reproducing organisms with short 
generation-time, esp. 
\dmel. However, there are many unresolved 
issues regarding computational/statistical methods:
\begin{itemize}
	\item Only a few tools can make full use of time-series points. Most just 
	handle	two time points. 
	\item As individuals of population are pooled and sequenced together, 
	coverage of a variant changes among different replicates and generations,
	which can have a critical effect on the power of different methods.
	\item A majority of existing tests, fail when the 
	initial 
	frequency of the favored allele is small, which is very likely in real word 
	scenarios.
	\item Except for multi-locus Gaussian process which is computationally 
	intractable, all the existing tools provide tests for single locus.
\end{itemize}



In this manuscript, we address these issues and develop a novel tool, \comale\
(available as open source) that can analyze time-series data, displays better 
power in detection of selection for a wide range of experimental evolution and 
selection parameters. We carefully examine the 
role of different experimental evolution parameters, including initial 
frequency of the favored allele, sampling time relative to the onset of 
selection, sequencing coverage, and the span of sampling on the power of 
selection.

We also found that existing tools for handling single time point data 
(often 
based on analysis of the site-frequency spectrum) and those for handling 
time-series data (based on a modeling of the favored allele frequency over 
time) are based on completely different principles, and no serious effort 
has been made to reconcile the two approaches.

In our experiments, we show that the problem of detecting a genomic region 
under selection is distinct from identifying the favored site that is 
functionally responding to selection constraint, and evolve-and-resequence 
strategies that work for detection may not be optimal for site-selection. In 
this sense, our analysis will guide future design of experimental evolution 
experiments. For all of these reasons, we hope that our manuscript will 
be of interest to 
\genetics\ readers. We thank you in advance for your consideration.

In terms of potential editors/reviewers, our work is of course most closely 
related to \genetics's Theoretical Population Genetics section.
In particular, Dr. Yun Song and Dr Rasmus Nielsen has previously published on 
the same topic.\\

Sincerely,


Arya Iranmehr

Vineet Bafna
\thispagestyle{empty}
\end{document}
