\documentclass[11pt]{article}
\title{\comale: Composition of Likelihoods for Evolve And Resequence 
Experiments}
\input{../manuscript/macros_arya.tex}
\date{\today}
\usepackage[margin=1in]{geometry} %not compatible with PLOS style
\renewcommand{\figurename}{Fig.}

\begin{document}
\noindent Dear Editor\\
\noindent PLoS Genetics\\

\noindent We would like to submit our manuscript  

``CLEAR: Composition of Likelihoods for Evolve And Resequence Experiments''

\noindent for possible publication in \emph{PLoS Genetics}. There are many 
reasons why 
the 
manuscript is appropriate for the journal. Experimental Evolution of organisms 
form an important methodology for understanding evolution in action. 
Specifically, they can help explain how organisms genetically adapt to 
selection pressures, e.g. pesticide, drug, antibiotic resistance, etc. With new 
sequencing technologies, we have the opportunity for ‘evolve-and-resequence’ 
experiments, providing a deep genomic sampling of the evolving population at 
different time points. \\

Increasingly, scientists are looking at sexually reproducing organisms, esp. 
\dmel, and acquiring time series data. However, there are many unresolved 
questions regarding computational/statistical methods that analyze the data in 
order to detect regions evolving under selection.
\begin{enumerate}
	\item Only a few tools can make full use of time-series points. Most just 
	handle 
	two time points. A recently published tool, Gaussian Process (GP), 
	published in 
	\emph{PLoS Genetics} 2015, can analyze time series through full likelihood 
	calculations, but its assumptions do not hold in real world scenarios. 
	Moreover, it is computationally intensive and does not scale to handling 
	genome 
	scale data. In our paper we deployed an improved model to improve the power.
	\item As individuals of population are pooled and sequenced together, 
	sequencing 
	coverage can have a critical effect on the power of different methods. We 
	evaluate different methods under various sequencing coverages and show that 
	CLEAR is robust to change of coverage.
	\item A majority of existing techniques, including GP, fail when the 
	initial 
	frequency of the favored allele is small, which is very likely in real word 
	scenarios. We show that CLEAR is not sensitive to this parameter and 
	performs 
	well over many different starting frequencies.
	\item We also found that existing tools for handling single time point data 
	(often 
	based on analysis of the site-frequency spectrum) and those for handling 
	time-series data (based on a modeling of the favored allele frequency over 
	time) are based on completely different principles, and no serious effort 
	has 
	been made to reconcile the two efforts.
\end{enumerate}


In this manuscript, we address these questions and develop a novel tool, CLEAR 
(available as open source) that can analyze time-series data, displays better 
power in detection of selection for a wide range of experimental evolution and 
selection parameters. Moreover, CLEAR is orders of magnitude faster than GP, 
the only other method that handles time-series data and estimates selection 
parameters.\\

It also resolves questions regarding the performance of site-frequency based 
methods by extending them to handle time-series data. We carefully examine the 
role of different experimental evolution parameters, including initial 
frequency of the favored allele, sampling time relative to the onset of 
selection, sequencing coverage, and the span of sampling on the power of 
selection. \\

In our experiments, we find that the problem of detecting a genomic region 
under selection is distinct from identifying the favored site that is 
functionally responding to selection constraint, and evolve-and-resequence 
strategies that work for detection may not be optimal for site-selection. In 
this sense, our calculations will guide future design of experimental evolution 
experiments.  As the methodology is the primary contribution, and understanding 
methods is necessary to understand results, we chose to put the methods section 
before the results. \\

For all of these reasons, we hope that our manuscript will be of interest to 
\emph{PLoS Genetics} readers. We thank you in advance for your consideration.\\

In terms of potential reviewers, our work is of course most closely related to 
Dr. Yun Song’s work which was published in \emph{PLoS Genetics} last year. We 
did 
discuss an early version of the manuscript with him, and even with competition, 
it would be OK to ask him to be a referee. We share one of the co-authors from 
that paper. Other potential reviewers include Dr. Dmitri Petrov who is doing a 
lot of work on experimental evolution, and some of the authors of the competing 
tools, including Hande Topa, Robert Kofler, Sergey Kryazhimskiy, and Alison 
Feder.\\ \\

Sincerely,\\


Arya Iranmehr

Vineet Bafna

\end{document}
